Online Classification of arXiv Papers Based on Your Preferences

### 2014-05-16 Paper.tex
small formatting edits in paper.

Let $I_1$ be the set of $t$ where a mistake was made and
$\widehat{y}=1$ (false positives).  Let $I_2$ be the set where a mistake was made with
$\widehat{y} = -1$ and $k = 1$ (false negatives where the algorithm updates), and $I_3$ the set where a mistake was
made with $\widehat{y} = -1$ and $k = 0$ (false negatives where the algorithm does {\it not} make an update). Finally, let $M_i = |I_i|$ for $i =
1,2,3$. We bound $M_1 + M_2$ in the same way we bound $M$ in the original Perceptron algorithm, because these are all the mistakes after which an update is made.. Assume there exists a $\rho$ and $R$ satisfying the assumptions of Theorem 7.8 from ~\cite{FML}, and a separating hyperplane determined by the vector $v$. Let $M' = M_1 + M_2$, then
\begin{align*}
M' \rho =&\ \frac{v}{\|v\|}\sum_{I_1 \cup I_2} y_tx_t \
 \leq   \Bigl\|\sum_{t \in I_1 \cup I_2} y_t x_t\Bigr\| \\
=& \ \Bigl\|\sum_{t \in I_1 \cup I_2} (w_{t+1} - w_t) \Bigr\| 
=  \|w_{T+1} \| \\
=&\ \sqrt{\sum_{t \in I_1 \cup I_2} \|  w_{t+1}\|^2 - \|w_t\|^2 }\\
= &\ \sqrt{\sum_{t \in I_1 \cup I_2} \|  w_{t} + y_t x_t\|^2 - \|w_t\|^2 }\\
= &\ \sqrt{\sum_{t \in I_1 \cup I_2} 2y_t w_t\cdot x_t - \|x_t\|^2} \\
\leq &\ \sqrt{\sum_{t \in I_1 \cup I_2} \|x_t\|^2}
\leq \sqrt{M'r^2}.
\end{align*}
This implies that $M' \leq \frac{r^2}{\rho^2}$. Since $M = M' + M_3$, a bound on $M_3$ is needed. We revert to a weaker bound and consider the expectation of
$M_3$. Since 
\[M_3 = \sum_{t=1}^T 1_{\left\{\widehat{y}_t = -1 \cap y_t = 1\right\}}1_{k_t = 0}, \]
we get that
\begin{align*}
\E[M_3] = & \sum_{t=1}^T \E[1_{\left\{\widehat{y_t} = -1 \cap y_t = 1\right\}}]\E[1_{k_t = 0}] \\
 = &\ (1 - p)\sum_{t=1}^T \E[1_{\left\{\widehat{y_t} = -1 \cap y_t = 1\right\}}],
\end{align*}
where the last equality uses the independence of $k_t$ from the data, and the expecation is taken over the realizations of $k_t$. To control the sum of expectations in this expression we consider the expectation of $M_2$. Since
\[
M_2 =  \sum_{t=1}^T 1_{\left\{\hat{y_t} = -1 \cap y_t = 1\right\}}1_{k_t = 1},
\]
then
\begin{align*}
\E[M_2] =&\ \sum_{t=1}^T \E[1_{\left\{\hat{y_t} = -1 \cap y_t = 1\right\}}]\E[1_{k_t = 1}] \\
=&\ p\sum_{t=1}^T \E[1_{\left\{\hat{y_t} = -1 \cap y_t = 1\right\}}].
\end{align*}
Thus
\[
E[M_3] = \ \frac{1-p}{p}\ \E[M_2].
\]
We can apply the crude bound $\E[M_3] \leq \frac{1-p}{p}\ \E[M']$, since $M_2 \leq M'$. Thus the expected number of mistakes can be bounded by
\begin{align*}
\E[M]  =  &\ \E[M'] + \E[M_3] \\
\leq & \ \frac{r^2}{\rho^2} + \frac{1-p}{p}\frac{r^2}{\rho^2} \\
\leq  & \ \frac{1}{p}\frac{r^2}{\rho^2}
\end{align*}
Note that this is a weak bound.  Not only is it greater than the
original perceptron bound by a factor of $\frac{1}{p}$, it is also only a
bound in expectation.

@@

\noindent Adaptive $p$ Perceptron algorithm:
\begin{quotation}
%\noindent Modified Perceptron:\\
\noindent $w_1=0$\\
\noindent $p=1$\\
{\tt for $t=1$ to $T$}:\\
\indent receive($x_t$)\\
\indent $\widehat{y}_t=\mbox{sign}(w_t\cdot x_t)$\\
\indent {\tt if} $\widehat{y}_t =1$ : \\
\indent\indent receive($y_t$)\\
\indent\indent {\tt if $y_t\neq\widehat{y}_t$}:\\
\indent\indent\indent $w_{t+1} = w_t + y_t x_t$\\
\indent\indent {\tt else $w_{t+1} = w_t$}\\
\indent {\tt else if ($\widehat{y}_t =-1$ {\tt and} $k_t = 1$)}:\\
\indent\indent {\tt if $y_t\neq\widehat{y}_t$}:\\
\indent\indent\indent $w_{t+1} = w_t + y_t x_t$\\
\indent\indent\indent $p=p + \gamma_1(1 - p)$\\
\indent\indent {\tt else }\\
\indent\indent\indent $w_{t+1} = w_t$\\
\indent\indent\indent $p = p - \gamma_2(p - p_{\min})$\\
\indent {\tt else if $\widehat{y}_t = -1$ and $k_t=0$}:\\
\indent\indent $w_{t+1} = w_t$\\
{\tt return $w_{T+1}$}
\end{quotation}
Note that $p_{\min}$, $\gamma_1$ and $\gamma_2$ are parameters that can be tuned. It is easy to extend the bound above from the modified Perceptron to the adaptive $p$ case.  Let $I_i$, $M_i$, and $M'$ be defined as above. The bound $M'\leq r^2/\rho^2$ follows in exactly the same way.  We then bound $M_3$ in expectation, 
\begin{align*}
\E[M_3] = & \sum_{t=1}^T \E[1_{\left\{\widehat{y_t} = -1 \cap y_t = 1\right\}}]\E[1_{k_t = 0}] \\
 \leq &\ (1 - p_{\min})\sum_{t=1}^T \E[1_{\left\{\widehat{y_t} = -1 \cap y_t = 1\right\}}],
\end{align*}
and bound $\E[M_2]$ from below
\begin{align*}
\E[M_2] =&\ \sum_{t=1}^T \E[1_{\left\{\hat{y_t} = -1 \cap y_t = 1\right\}}]\E[1_{k_t = 1}] \\
\geq &\ p_{\min}\sum_{t=1}^T \E[1_{\left\{\hat{y_t} = -1 \cap y_t = 1\right\}}].
\end{align*}
Continuing as before, this gives the following bound on $\E[M]$,
\begin{align*}
\E[M] = &\ \E[M'] + \E[M_3] \leq \frac{r^2}{\rho^2} + \frac{1-p_{\min}}{p_{\min}} \frac{r^2}{\rho^2}\\
= &\ \frac{1}{p_{\min}}\frac{r^2}{\rho^2}
\end{align*}
Unfortunately, we do not have a stonger theoretical bound for this adaptive $p$ approach, but we test this scheme in section \ref{sec:results} and see that it performs fairly well while requiring less input from the user.

@@

\begin{figure}[h!]\centering
\includegraphics[width=0.65\textwidth]{EveryoneRecallPaper.pdf}
\includegraphics[width=0.65\textwidth]{EveryonePrecisionPaper.pdf}
\label{fig:PrecisionandRecall}
\caption{Performance of tested online classification algorithms using different values of $p$.  The adaptive $p$ algorithm performs relatively well compared to the original algorithm ($p = 1$), and requires less than half as many negative updates. The $p = 0.1$ approach has very poor recall, but requires far fewer updates.  The majority vote classifier combines all others in this figure.}
\end{figure}

>>>>

Let $I_1$ be the set of $t$ where a mistake was made and
$\widehat{y}=1$ (false positives).  Let $I_2$ be the set where a mistake was made with
$\widehat{y} = -1$ and $k = 1$ (false negatives where the algorithm updates), and $I_3$ the set where a mistake was
made with $\widehat{y} = -1$ and $k = 0$ (false negatives where the algorithm does {\it not} make an update). Finally, let $M_i = |I_i|$ for $i =
1,2,3$. We bound $M_1 + M_2$ in the same way we bound $M$ in the original Perceptron algorithm, because these are all the mistakes after which an update is made.. Assume there exists a $\rho$ and $R$ satisfying the assumptions of Theorem 7.8 from ~\cite{FML}, and a separating hyperplane determined by the vector $v$. Let $M' = M_1 + M_2$, then
\begin{align*}
M' \rho =&\ \frac{v}{\|v\|}\sum_{I_1 \cup I_2} y_tx_t  \leq \Bigl\|\sum_{t \in I_1 \cup I_2} y_t x_t\Bigr\| \\
=& \ \Bigl\|\sum_{t \in I_1 \cup I_2} (w_{t+1} - w_t) \Bigr\| =  \|w_{T+1} \| \\
=&\ \sqrt{\sum_{t \in I_1 \cup I_2} \|  w_{t+1}\|^2 - \|w_t\|^2 }\\
= &\ \sqrt{\sum_{t \in I_1 \cup I_2} \|  w_{t} + y_t x_t\|^2 - \|w_t\|^2 }\\
= &\ \sqrt{\sum_{t \in I_1 \cup I_2} 2y_t w_t\cdot x_t - \|x_t\|^2} \\
\leq &\ \sqrt{\sum_{t \in I_1 \cup I_2} \|x_t\|^2} \leq \sqrt{M'r^2}.
\end{align*}
This implies that $M' \leq \frac{r^2}{\rho^2}$. Since $M = M' + M_3$, a bound on $M_3$ is needed. We revert to a weaker bound and consider the expectation of
$M_3$. Since 
\[M_3 = \sum_{t=1}^T 1_{\left\{\widehat{y}_t = -1 \cap y_t = 1\right\}}1_{k_t = 0}, \]
we get that
\begin{align*}
\E[M_3] = & \sum_{t=1}^T \E[1_{\left\{\widehat{y_t} = -1 \cap y_t = 1\right\}}]\E[1_{k_t = 0}] = (1 - p)\sum_{t=1}^T \E[1_{\left\{\widehat{y_t} = -1 \cap y_t = 1\right\}}],
\end{align*}
where the last equality uses the independence of $k_t$ from the data, and the expecation is taken over the realizations of $k_t$. To control the sum of expectations in this expression we consider the expectation of $M_2$. Since
\[
M_2 =  \sum_{t=1}^T 1_{\left\{\hat{y_t} = -1 \cap y_t = 1\right\}}1_{k_t = 1},
\]
then
\begin{align*}
\E[M_2] =&\ \sum_{t=1}^T \E[1_{\left\{\hat{y_t} = -1 \cap y_t = 1\right\}}]\E[1_{k_t = 1}] = p\sum_{t=1}^T \E[1_{\left\{\hat{y_t} = -1 \cap y_t = 1\right\}}].
\end{align*}
Thus
\[
E[M_3] = \ \frac{1-p}{p}\ \E[M_2].
\]
We can apply the crude bound $\E[M_3] \leq \frac{1-p}{p}\ \E[M']$, since $M_2 \leq M'$. Thus the expected number of mistakes can be bounded by
\begin{align*}
\E[M]  =  &\ \E[M'] + \E[M_3] \\
\leq & \ \frac{r^2}{\rho^2} + \frac{1-p}{p}\frac{r^2}{\rho^2} \\
\leq  & \ \frac{1}{p}\frac{r^2}{\rho^2}
\end{align*}
Note that this is a weak bound.  Not only is it greater than the
original perceptron bound by a factor of $\frac{1}{p}$, it is also only a
bound in expectation.

@@

\noindent Adaptive $p$ Perceptron algorithm:
\begin{quotation}
%\noindent Modified Perceptron:\\
\noindent $w_1=0$\\
\noindent $p=1$\\
{\tt for $t=1$ to $T$}:\\
\indent receive($x_t$)\\
\indent $\widehat{y}_t=\mbox{sign}(w_t\cdot x_t)$\\
\indent {\tt if} $\widehat{y}_t =1$ : \\
\indent\indent receive($y_t$)\\
\indent\indent {\tt if $y_t\neq\widehat{y}_t$}:\\
\indent\indent\indent $w_{t+1} = w_t + y_t x_t$\\
\indent\indent {\tt else $w_{t+1} = w_t$}\\
\indent {\tt else if ($\widehat{y}_t =-1$ {\tt and} $k_t = 1$)}:\\
\indent\indent {\tt if $y_t\neq\widehat{y}_t$}:\\
\indent\indent\indent $w_{t+1} = w_t + y_t x_t$\\
\indent\indent\indent $p=p + \gamma_1(1 - p)$\\
\indent\indent {\tt else }\\
\indent\indent\indent $w_{t+1} = w_t$\\
\indent\indent\indent $p = p - \gamma_2(p - p_{\min})$\\
\indent {\tt else if $\widehat{y}_t = -1$ and $k_t=0$}:\\
\indent\indent $w_{t+1} = w_t$\\
{\tt return $w_{T+1}$}
\end{quotation}
Note that $p_{\min}$, $\gamma_1$ and $\gamma_2$ are parameters that can be tuned. It is easy to extend the bound above from the modified Perceptron to the adaptive $p$ case.  Let $I_i$, $M_i$, and $M'$ be defined as above. The bound $M'\leq r^2/\rho^2$ follows in exactly the same way.  We then bound $M_3$ in expectation, 
\begin{align*}
\E[M_3] = & \sum_{t=1}^T \E[1_{\left\{\widehat{y_t} = -1 \cap y_t = 1\right\}}]\E[1_{k_t = 0}] \leq \  (1 - p_{\min})\sum_{t=1}^T \E[1_{\left\{\widehat{y_t} = -1 \cap y_t = 1\right\}}],
\end{align*}
and bound $\E[M_2]$ from below
\begin{align*}
\E[M_2] =&\ \sum_{t=1}^T \E[1_{\left\{\hat{y_t} = -1 \cap y_t = 1\right\}}]\E[1_{k_t = 1}] \geq \ p_{\min}\sum_{t=1}^T \E[1_{\left\{\hat{y_t} = -1 \cap y_t = 1\right\}}].
\end{align*}
Continuing as before, this gives the following bound on $\E[M]$,
\begin{align*}
\E[M] = &\ \E[M'] + \E[M_3] \leq \frac{r^2}{\rho^2} + \frac{1-p_{\min}}{p_{\min}} \frac{r^2}{\rho^2}\\
= &\ \frac{1}{p_{\min}}\frac{r^2}{\rho^2}
\end{align*}
Unfortunately, we do not have a stronger theoretical bound for this adaptive $p$ approach that takes into account the fact that $p$ will become smaller as the algorithm learns. However, we test this scheme in section \ref{sec:results} and see that it performs fairly well while requiring less input from the user.

@@

\begin{figure}[h!]\centering
\includegraphics[width=0.65\textwidth]{EveryoneRecallPaper.pdf}
\includegraphics[width=0.65\textwidth]{EveryonePrecisionPaper.pdf}
\label{fig:PrecisionandRecall}
\caption{ \small Performance of tested online classification algorithms using different values of $p$.  The adaptive $p$ algorithm performs relatively well compared to the original algorithm ($p = 1$), and requires less than half as many negative updates. The $p = 0.1$ approach has very poor recall, but requires far fewer updates.  The majority vote classifier combines all others in this figure.}
\end{figure}
